---
title: 'DOSPERT Data: combined financial scale'
author: "Salah Chafik <salah@decisionsciences.columbia.edu>"
date: "Updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: yes
    self_contained: no
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("/git_repositories/bment/scripts/helper_scripts/plotting_functions.R")
source("/git_repositories/bment/scripts/helper_scripts/load_packages.R")
#be sure to point the below to where you source the data
full_data <- readRDS("/git_repositories/betterment/data/rds_data/dospert_data.RDS")
```

#Risk Taking
####higher number = greater risk taking 
  + 42 is the max (6 questions, likert scale of 7 options)
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "finRT_sum") + ggtitle("risk taking") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$finRT_sum)
```

\
[back to table of contents](#TOC)

#Risk Benefit
####higher number = greater benefits
  + 42 is the max (6 questions, likert scale of 7 options)
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "finRB_sum") + ggtitle("risk benefit") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$finRB_sum)
```

\
[back to table of contents](#TOC)

#Risk Perception
####higher number = perceptions of greater risk 
  + 42 is the max (6 questions, likert scale of 7 options)
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "finRP_sum") + ggtitle("risk perception") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$finRP_sum)
```

\
[back to table of contents](#TOC)

#Risk Seeking/Aversion: beta
####Positive coef(b) = **risk seeking** behavior, Negative coef(b) = **risk aversion** behavior
  + Preference (X) = a(Expected Benefit(X)) + b(Perceived Risk(X)) + c
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "fin_RP") + ggtitle("all participants") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$fin_RP)
```

> participants are slightly risk averse overall

\
[back to table of contents](#TOC)

#Dohmen

  + "How do you see yourself? Are you generally a person who is fully prepared to take risks or avoid them?"
    + **0 = Not at all willing to take risks**
    + **10 = Very willing to take risks**
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "dohmen") + scale_x_continuous(breaks=seq(0, 10, 1))
summary(full_data$dohmen)
```

> people perceive themselves to be on average slightly more willing to take risks

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "fin_RP") + ggtitle("all participants") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$fin_RP)
```
\
[back to table of contents](#TOC)

```{r, echo=FALSE, eval=FALSE}

#General Risk Taking Propensity 
####Positive value means people have a general propensity to take risk
  + Preference (X) = a(Expected Benefit(X)) + b(Perceived Risk(X)) + c
    + intercept from above model  
    
#```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(full_data, "fin_int") + ggtitle("risk taking propensity") + theme(text=element_text(family="Georgia", face="bold"))
summary(full_data$fin_int)
#```

####takeaway 
> on average people have slight propensity to take risk

\
[back to table of contents](#TOC)
```


#Correlations: SPLOM plots
####risk taking, risk benefit and risk perception scores
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, cache=TRUE}
yo <- select(full_data, risk_taking = finRT_sum, 
             risk_perception = finRP_sum, risk_benefit = finRB_sum, dohmen)
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Risk Taking/Benefits/Perception SPLOM Plot")  
```

####takeaway 
> risk perception and risk taking are weakly negatively correlated (-.26) i.e. the more risk someone perceives the less willing s/he is to take a risk 

> risk benefit and risk taking are moderately positively correlated (.33) i.e. the more beneficial someone perceives a risk, the more willing s/he is is to take that risk

> dohmen and risk taking are moderately positively correlated (.36) i.e. the more risk you are willing to take the more you see yourself as a person willing to take risks 

> dohmen and risk benefit are moderately positively correlated (.32) i.e. the more benefit you see in a risk the more you see yourself as a person willing to take risks 

\
[back to table of contents](#TOC)

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, cache=TRUE, eval=FALSE}
##coefficients
  + **risk taking propensity** (incercept) **reward sensitivity** (alpha) and **risk seeking/aversion** (beta) scores

yo <- select(full_data, propensity = fin_int, 
             alpha = fin_RB, beta = fin_RP)
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Coefficients SPLOM Plot") 

####takeaway
> alpha and beta are strongly positively correlated (.60) i.e. the more sensitive someone is to reward the more risk seeking s/he is

> alpha and propensity to take risk are very strongly negatively correlated (-.87) i.e. the more reward sensitive a person is, the less someone's propensity to take a risk

> beta and propensity to take risk are very strongly negatively correlated (-.89) i.e. the more risk seeking behavior someone exhibits, the less someone's propensityv to take a risk

\
[back to table of contents](#TOC)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=10, cache=TRUE, eval=FALSE}
####everything
yo <- full_data %>%
 dplyr::select(risk_taking = finRT_sum, risk_benefit = finRB_sum, 
               risk_perception = finRP_sum, propensity = fin_int, 
               alpha = fin_RB, beta = fin_RP) 
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "All DOSPERT Variables SPLOM Plot") 
\
[back to table of contents](#TOC)
```  




