---
title: 'DOSPERT Data: Gambling vs Investment'
author: "Salah Chafik <salah@decisionsciences.columbia.edu>"
date: "Updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: yes
    self_contained: no
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("/git_repositories/bment/scripts/helper_scripts/plotting_functions.R")
source("/git_repositories/bment/scripts/helper_scripts/load_packages.R")
#be sure to point the below to where you source the data
full_data <- readRDS("/git_repositories/betterment/data/rds_data/dospert_data.RDS") %>%
              select(unique_id, dohmen)
gamble <- readRDS("/git_repositories/betterment/data/rds_data/dospert_gamble.RDS")
gamble <- left_join(gamble, full_data, by = "unique_id")
invest <- readRDS("/git_repositories/betterment/data/rds_data/dospert_invest.RDS")
invest <- left_join(invest, full_data, by = "unique_id")
```

#Risk Taking
####higher number = greater risk taking 
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(gamble, "finRT_sum_G") + ggtitle("gambling domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(gamble$finRT_sum_G)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(invest, "finRT_sum_I") + ggtitle("investment domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(invest$finRT_sum_I)
```

####takeaway 
> people are quite willing to take investment risks, and very unwilling to take gambling risks

\
[back to table of contents](#TOC)

#Risk Benefit
####higher number = greater benefits
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(gamble, "finRB_sum_G") + ggtitle("gambling domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(gamble$finRB_sum)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(invest, "finRB_sum_I") + ggtitle("investment domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(invest$finRB_sum_I)
```

####takeaway 
> people find more benefit from risk in investment domain over gambling domain (by a factor of 2)

\
[back to table of contents](#TOC)

#Risk Perception
####higher number = perceptions of greater risk 
  
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(gamble, "finRP_sum_G") + ggtitle("gambling domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(gamble$finRP_sum_G)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(invest, "finRP_sum_I") + ggtitle("investment domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(invest$finRP_sum_I)
```

####takeaway
> people feel that there is more risk in gambling domain vs investment domain

\
[back to table of contents](#TOC)

```{r, eval=FALSE, echo=FALSE}
#Risk Seeking/Aversion: beta
####Positive coef(b) = **risk seeking** behavior, Negative coef(b) = **risk aversion** behavior
  + Preference (X) = a(Expected Benefit(X)) + b(Perceived Risk(X)) + c
#```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(gamble, "fin_RP_G") + ggtitle("gambling domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(gamble$fin_RP_G)
#```

#```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(invest, "fin_RP_I") + ggtitle("investment domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(invest$fin_RP_I)
#```

####takeaway 
> participants are slightly more risk averse in investment domain

\
[back to table of contents](#TOC)
```

```{r, eval=FALSE, echo=FALSE}
#General Risk Taking Propensity 
####Positive value means people have a propensity to take risk
  + lm(risk taking propensity ~ benefit + perception)
    + intercept from above model
  
#```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(gamble, "fin_int_G") + ggtitle("gambling domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(gamble$fin_int_G)
#```

#```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, cache=TRUE}
plot_hist(invest, "fin_int_I") + ggtitle("investment domain") + theme(text=element_text(family="Georgia", face="bold"))
summary(invest$fin_int_I)
#```

####takeaway
> people have slightly higher propensity to take risk in investment domain vs. gambling domain

\
[back to table of contents](#TOC)
```


#Correlations: SPLOM plots
```{r, echo=FALSE}
#   + reminder when r = 
#     + .70 or higher  Very strong positive relationship 
#     + .40 to +.69  Strong positive relationship 
#     + .30 to +.39	Moderate positive relationship 
#     + .20 to +.29	weak positive relationship 
#     + .01 to +.19	No or negligible relationship 
#     + -.01 to -.19	No or negligible relationship 
#     + -.20 to -.29	weak negative relationship 
#     + -.30 to -.39	Moderate negative relationship 
#     + -.40 to -.69	Strong negative relationship 
#     + -.70 or higher	Very strong negative relationship
```
##gambling
#####risk taking, risk benefit and risk perception scores
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=7, cache=TRUE}
yo <- left_join(gamble, invest, by = "unique_id") %>%
        select(risk_taking = finRT_sum_G, risk_benefit = finRB_sum_G, 
               risk_perception = finRP_sum_G, dohmen = dohmen.x)
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Gambling: Risk Taking/Benefits/Perception SPLOM Plot") 
```

####takeaway

> risk perception and risk taking are moderately negatively correlated (-.36) i.e. the more risky someone perceives a risk, the less willing s/he is to take that risk

> risk perception and risk benefit are weakly negatively correlated (-.21) i.e. the more risky someone perceives a risk, the less beneficial a s/he perceives that risk

> risk benefit and risk taking are weakly positively correlated (.26) i.e. the more beneficial someone perceives a risk, the more willing s/he is to take that risk

> perceived risks are a stronger predictor for **gambling** risk taking 

##investment
#####risk taking, risk benefit and risk perception scores
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=7, cache=TRUE}
yo <- left_join(gamble, invest, by = "unique_id") %>%
        select(risk_taking = finRT_sum_I, risk_benefit = finRB_sum_I, 
               risk_perception = finRP_sum_I, dohmen = dohmen.x)
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Investment: Risk Taking/Benefits/Perception SPLOM Plot") 
```


####takeaway

> risk perception and risk taking are weakly negatively correlated (-.18) i.e. the more risky someone perceives a risk, the less willing s/he is to take that risk 

> risk perception and risk benefit are weakly negatively correlated (-.18) i.e. the more risky someone perceives a risk, the less beneficial s/he perceives that risk

> risk benefit and risk taking are moderately positively correlated (.40) i.e. the more willing someone is to take a risk the more beneficial s/he perceives that risk  

> perceived benefits are a stronger predictor for **investment** risk taking.  

\
[back to table of contents](#TOC)

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=7, cache=TRUE, eval=FALSE}
####gambling vs investment
#####risk taking, risk benefit and risk perception scores
yo <- left_join(gamble, invest, by = "unique_id") %>%
        select(risk_taking_I = finRT_sum_I, risk_taking_G = finRT_sum_G, 
               risk_benefit_I = finRB_sum_I, risk_benefit_G = finRB_sum_G, 
               risk_perception_I = finRP_sum_I, risk_perception_G = finRP_sum_G) 
ggpairs(yo, lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Gamble vs Invest: Risk Taking/Benefits/Perception SPLOM Plot") 
```

