---
title: 'Predicting Allocation Deviation'
author: "Salah Chafik <salah@decisionsciences.columbia.edu>"
date: "Updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: yes
    self_contained: no
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("/git_repositories/bment/scripts/helper_scripts/plotting_functions.R")
source("/git_repositories/bment/scripts/helper_scripts/load_packages.R")
#be sure to point the below to where you source the data
full_data <- readRDS("/git_repositories/betterment/data/rds_data/userANDdospert_full.RDS") %>%
              mutate(goal_type = factor(goal_type)) %>% #making sure goal type is a factor
              rename(DOSPERT_riskTaking = finRT_sum, beta = fin_RP) #renaming for clarity
full_data$goal_type <- relevel(full_data$goal_type, ref = "EMERGENCY") #setting emergency as baseline
gamble <- readRDS("/git_repositories/betterment/data/rds_data/dospert_gamble.RDS")
invest <- readRDS("/git_repositories/betterment/data/rds_data/dospert_invest.RDS")
gamble <- left_join(full_data, gamble %>% select(finRT_sum_G:fin_RP_G, unique_id), 
                    by=c("visitor_key" = "unique_id"))
invest <- left_join(full_data, invest %>% select(finRT_sum_I:fin_RP_I, unique_id), 
                    by=c("visitor_key" = "unique_id"))
```


#OLS regression
#####using deviation (between actual and recommended allocation) as the dependent variable, we examine three significant explanatory variables: DOSPERT hypothetical risk taking, DOSPERT beta score (risk attitude), and Dohmen self reported risk

```{r, echo=FALSE, results = 'asis'}
# summary(lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)) 
# summary(lm(allocation_deviation ~ beta + allocation_stock_rec, data = full_data))
# summary(lm(allocation_deviation ~ dohmen + allocation_stock_rec, data = full_data))
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)
model2 <- lm(allocation_deviation ~ beta + allocation_stock_rec, data = full_data)
model3 <- lm(allocation_deviation ~ dohmen + allocation_stock_rec, data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: controling for recommendation',
        caption.above = TRUE, digits = 4)
```

  + **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + recommended allocation)

> **for a one unit increase in DOSPERT hypothetical risk taking, there is a .58 percentage point increase in user allocation deviation (i.e. an increase in stock allocation above the recommended amount). **

  + **model 2**
    + lm(deviation ~ **DOSPERT beta** + recommended allocation)

> **for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 2 percentage point increase in user allocation deviation**

  + **model 3**
    + lm(deviation ~ **dohmen** + recommended allocation)

> **for a one unit increase in dohmen self reported risk, there is a 1.31 percentage point increase in user allocation deviation**

####scatterplot matrix (SPLOM) plot
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=8, cache=TRUE}
ggpairs(full_data %>% select(allocation_deviation, DOSPERT_riskTaking, beta, dohmen),
        lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Allocation Deviation vs Hypothetical Risk Taking/Risk Attitude(Beta)/Dohmen") 
```

> no significant correlations between allocation deviation and three explanatory variables

\
[back to table of contents](#TOC)

#OLS: by goal type
 
####[as seen previously](https://vlab.decisionsciences.columbia.edu/salah/user_goal_ii.html#deviations), goal type is significant as a predictive variable:
```{r, echo=FALSE}
summary(aov(allocation_deviation ~ goal_type, data=full_data))
```

```{r, echo=FALSE, results = 'asis'}
# summary(lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec + goal_type, data = full_data)) 
# summary(lm(allocation_deviation ~ beta + allocation_stock_rec + goal_type, data = full_data))
# summary(lm(allocation_deviation ~ dohmen + allocation_stock_rec + goal_type, data = full_data))
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec + goal_type, data = full_data)
model2 <- lm(allocation_deviation ~ beta + allocation_stock_rec + goal_type, data = full_data)
model3 <- lm(allocation_deviation ~ dohmen + allocation_stock_rec + goal_type, data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by goal type',
        caption.above = TRUE, digits = 4)
```

> **while users generally allocate more stocks than advised, they do more so for `education`, `house`, `purchase`, and `retirement` goal types, relative to the emergency goal type baseline. this is true of our 3 risk taking predictors (DOSPERT risk taking, Dospert beta and dohmen) when controlling for recommended allocation. ** 

\
[back to table of contents](#TOC)

#OLS: by DOSPERT sub-scales
```{r, echo=FALSE, results='asis'}
ss <- left_join(gamble, invest %>% select(account_id, DOSPERT_riskTaking_I = finRT_sum_I, 
                                          beta_I = fin_RP_I), by = "account_id") %>%
              rename(DOSPERT_riskTaking_G = finRT_sum_G, beta_G = fin_RP_G)
model1_ss <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + 
                                       allocation_stock_rec, data = ss)

htmlreg(l = list(model1_ss), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by DOSPERT sub-scales',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking: gamble** + **DOSPERT risk taking: invest** + recommended allocation)

> **investment is the more predictive portion of the DOSPERT scale.**

\
[back to table of contents](#TOC)

#OLS: by Dohmen
```{r, echo=FALSE, results='asis'}
model1_d <- lm(allocation_deviation ~ DOSPERT_riskTaking + dohmen + allocation_stock_rec, data = full_data)
model2_d <- lm(allocation_deviation ~ beta +  dohmen + allocation_stock_rec, data = full_data)

htmlreg(l = list(model1_d, model2_d), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by Dohmen',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + dohmen + recommended allocation)

> no difference after inclusion of dohmen 

+ **model 2**
    + lm(deviation ~ **DOSPERT beta** + dohmen + recommended allocation)

> estimate is smaller and p-val is higher (.059 to .078) after inclusion of dohmen 

\
[back to table of contents](#TOC)

#OLS: by DOSPERT sub-scales + Dohmen

```{r, echo=FALSE, results='asis'}
model1_dss <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + 
                                        dohmen + allocation_stock_rec, data = ss)

htmlreg(l = list(model1_dss), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by DOSPERT sub-scales + Dohmen',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking: gamble** + **DOSPERT risk taking: invest** + dohmen + recommended allocation)

> effect size decreased and p-value increased (.001 to .008) by inclusion of dohmen

\
[back to table of contents](#TOC)

#OLS: combining risk taking predictors

```{r, echo=FALSE, results='asis'}
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec, data = full_data) 
model2 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type, data = full_data) 
model3 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + DOSPERT_riskTaking*goal_type, 
                                    data = full_data) #Drt
model4 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + beta*goal_type, 
                                    data = full_data) #beta
model5 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + dohmen*goal_type, 
                                    data = full_data) #dohmen
htmlreg(l = list(model1, model2, model3, model4, model5), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: Combining Risk Taking Predictors',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + **DOSPERT beta** + **dohmen** + recommended allocation)

> **for a one unit increase in DOSPERT hypothetical risk taking, there is a .39 percentage point increase in user allocation deviation (i.e. an increase in stock allocation above the recommended amount).**

+ **model 2**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + **goal type** + recommended allocation)

> **with the inclusion of goal type in the model, the same significant relationship holds true for DOSPERT hypothetical risk taking as in Model 1. in addition, for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 1.74 percentage point increase in user allocation deviation. finally, users allocate more stocks for `education`, `house`, `purchase`, and `retirement` goal types, relative to the emergency goal baseline.**

+ **model 3**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **DOSPERT risk taking**X**goal_type**  + recommended allocation)
  
> with the inclusion of Dospert risk taking and goal type interaction, the same significant relationship holds true for DOSPERT beta as in Model 2. 

+ **model 4**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **DOSPERT beta**X**goal_type**  + recommended allocation)

> **with the inclusion of Dospert beta and goal type interaction, the same significant relationship holds true for DOSPERT risk taking as in Models 1 + 2. in addition, for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 8.26/10.09 percentage point increase in user allocation deviation for `house`/`purchase` goal types.** finally, users allocate more stocks for `house` and `purchase` goal types, relative to the emergency goal baseline.

+ **model 5**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **dohmen**X**goal_type**  + recommended allocation)

> with the inclusion of dohmen and goal type interaction, the same significant relationship holds true for DOSPERT risk taking as in Models 1 + 2 + 4. the same significant holds true for DOSPERT beta from Models 2 + 3. in addition, users allocate more stocks for `education` goal type, relative to the emergency goal baseline. finally, for a one unit increase in dohmen, there is a 4.96 percentage point decrease in user allocation deviation for `education` goal type.

\
[back to table of contents](#TOC)


#logit regression
```{r, echo=FALSE, warning=FALSE, message=FALSE}
full_data <- full_data %>% mutate(deviation = ifelse(is.na(allocation_deviation), NA,
                                  ifelse(allocation_deviation > 0, "positive",
                                          ifelse(allocation_deviation == 0, "exact","negative"))))

print("deviation broken down by groups")
table(full_data$deviation)
```


```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
library(mlogit)

wide_full <- mlogit.data(full_data, choice = "deviation", shape = "wide")

model1 <- mlogit(deviation ~ 1 | DOSPERT_riskTaking + allocation_stock_rec, data = wide_full, reflevel = "exact")
model2 <- mlogit(deviation ~ 1 | beta + allocation_stock_rec, data = wide_full, reflevel = "exact")
model3 <- mlogit(deviation ~ 1 | dohmen + allocation_stock_rec, data = wide_full, reflevel = "exact")

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Deviation: multinomial logistic',
        caption.above = TRUE, digits = 4)

# model1 <- mlogit(deviation ~ 1 | DOSPERT_riskTaking + allocation_stock_rec, data = wide_full, reflevel = "negative")
# model2 <- mlogit(deviation ~ 1 | beta + allocation_stock_rec, data = wide_full, reflevel = "negative")
# model3 <- mlogit(deviation ~ 1 | dohmen + allocation_stock_rec, data = wide_full, reflevel = "negative")
#
#
# htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
#         bold = 0.1, caption = 'Deviation: multinomial logistic (ref = "negative")',
#         caption.above = TRUE, digits = 4)
#
#
# model1 <- mlogit(deviation ~ 1 | DOSPERT_riskTaking + allocation_stock_rec, data = wide_full, reflevel = "positive")
# model2 <- mlogit(deviation ~ 1 | beta + allocation_stock_rec, data = wide_full, reflevel = "positive")
# model3 <- mlogit(deviation ~ 1 | dohmen + allocation_stock_rec, data = wide_full, reflevel = "positive")
#
#
# htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
#         bold = 0.1, caption = 'Deviation: multinomial logistic (ref = "positive")',
#         caption.above = TRUE, digits = 4)

```

> the higher DOSPERT risk taking score/DOSPERT beta/Dohmen is, the more likely that the person will deviate positively (compared to following the recommended allocation amount).

> however, none of our variables are significantly predictive of negative deviation. that is to say, even when the DOSPERT risk taking score is low, they are not more likely to deviate negatively. 

> the higher the recommended allocation, the less likely that the person will positively deviate, compared to following the recommendation. but again, allocation recommendation is not significant for comparison between following the recommendation and negative deviation.

\
[back to table of contents](#TOC)

#logit: by current balance
```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}

model1 <- mlogit(deviation ~ 1 | DOSPERT_riskTaking + allocation_stock_rec + user_current_balance, data = wide_full, reflevel = "exact")
model2 <- mlogit(deviation ~ 1 | beta + allocation_stock_rec + user_current_balance, data = wide_full, reflevel = "exact")
model3 <- mlogit(deviation ~ 1 | dohmen + allocation_stock_rec + user_current_balance, data = wide_full, reflevel = "exact")

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Deviation: multinomial logistic including current balance',
        caption.above = TRUE, digits = 4)
```

> **people are more likely to follow the recommendation when their current balance is higher.**

\
[back to table of contents](#TOC)

```{r, echo=FALSE, results='asis', eval=FALSE}
#Allocation Deviation: combining risk taking predictors (sub-scales)
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec, data = ss) 
model2 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + 
                                    DOSPERT_riskTaking_G*goal_type + DOSPERT_riskTaking_I*goal_type, 
                                    data = ss) #Drt
model3 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + beta*goal_type, 
                                    data = ss) #beta
model4 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + dohmen*goal_type, 
                                    data = ss) #dohmen
htmlreg(l = list(model1, model2, model3, model4), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: Combining Risk Taking Predictors (DOSPERT sub-scales)',
        caption.above = TRUE, digits = 4)
\
[back to table of contents](#TOC)
```



```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=7, eval=FALSE}
#Actual Allocation? 
ggpairs(full_data %>% select(allocation_stock_actual, DOSPERT_riskTaking),
        lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Actual Stock Allocation vs Hypothetical Risk Taking SPLOM Plot") 

> there is a weak (although strongly significant) positive correlation of .11 between actual stock allocation and DOSPERT hypothetical risk taking 

```




