
---
title: 'Predicting Allocation'
author: "Salah Chafik <salah@decisionsciences.columbia.edu>"
date: "Updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: yes
    self_contained: no
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("/git_repositories/bment/scripts/helper_scripts/plotting_functions.R")
source("/git_repositories/bment/scripts/helper_scripts/load_packages.R")
#be sure to point the below to where you source the data
full_data <- readRDS("/git_repositories/betterment/data/rds_data/userANDdospert_full.RDS") %>%
              mutate(goal_type = factor(goal_type)) %>% #making sure goal type is a factor
              rename(DOSPERT_riskTaking = finRT_sum, beta = fin_RP) #renaming for clarity
full_data$goal_type <- relevel(full_data$goal_type, ref = "RETIREMENT") #setting retirement as baseline
gamble <- readRDS("/git_repositories/betterment/data/rds_data/dospert_gamble.RDS")
invest <- readRDS("/git_repositories/betterment/data/rds_data/dospert_invest.RDS")
gamble <- left_join(full_data, gamble %>% select(finRT_sum_G:fin_RP_G, unique_id), 
                    by=c("visitor_key" = "unique_id"))
invest <- left_join(full_data, invest %>% select(finRT_sum_I:fin_RP_I, unique_id), 
                    by=c("visitor_key" = "unique_id"))
```

#OLS regression
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
model1 <- lm(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)
model2 <- lm(allocation_stock_actual ~ beta + allocation_stock_rec, data = full_data)
model3 <- lm(allocation_stock_actual ~ dohmen + allocation_stock_rec, data = full_data)

# model1 <- truncreg(allocation_stock_actual ~ DOSPERT_riskTaking, data = full_data, 
#                    point = 1, direction = "right")
# model2 <- truncreg(allocation_stock_actual ~ beta, data = full_data, 
#                    point = 1, direction = "right")
# model3 <- truncreg(allocation_stock_actual ~ dohmen, data = full_data, 
#                    point = 1, direction = "right")

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation: OLS Regression',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(allocation ~ **DOSPERT risk taking** + recommended allocation)

> for a one unit increase in DOSPERT hypothetical risk taking, there is a .58 percentage point increase in user stock allocation 


  + **model 2**
    + lm(allocation ~ **DOSPERT beta** + recommended allocation)

> for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 2 percentage point increase in user allocation

  + **model 3**
    + lm(allocation ~ **dohmen** + recommended allocation)

> for a one unit increase in dohmen self reported risk, there is a 1.31 percentage point increase in user allocation


#truncated regression

> this technique is utilized when data for the outcome variable is only available for part of the sample (i.e. truncated sample). this is likely not the case for our study.

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(truncreg) # http://www.ats.ucla.edu/stat/r/dae/truncreg.htm

model1 <- truncreg(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data, 
                   point = 1, direction = "right")
model2 <- truncreg(allocation_stock_actual ~ beta + allocation_stock_rec, data = full_data, 
                   point = 1, direction = "right")
model3 <- truncreg(allocation_stock_actual ~ dohmen + allocation_stock_rec, data = full_data, 
                   point = 1, direction = "right")

#not controlling for rec
# model1 <- truncreg(allocation_stock_actual ~ DOSPERT_riskTaking, data = full_data, 
#                    point = 1, direction = "right")
# model2 <- truncreg(allocation_stock_actual ~ beta, data = full_data, 
#                    point = 1, direction = "right")
# model3 <- truncreg(allocation_stock_actual ~ dohmen, data = full_data, 
#                    point = 1, direction = "right")

#hack for displaying in a neat table
full_data$sigma <- sample(c(0,1), 646, replace = TRUE)
pseudo1 <- lm(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec + sigma, data = full_data)
pseudo2 <- lm(allocation_stock_actual ~ beta + allocation_stock_rec + sigma, data = full_data)
pseudo3 <- lm(allocation_stock_actual ~ dohmen + allocation_stock_rec + sigma, data = full_data)

htmlreg(l = list(pseudo1, pseudo2, pseudo3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        override.coef = list(model1$coefficients, model2$coefficients, model3$coefficients),
        override.se = list(sqrt(diag(model1$vcov)),sqrt(diag(model2$vcov)),sqrt(diag(model3$vcov))),
        override.pval = list(summary(model1)$coefficients[,4],
                             summary(model2)$coefficients[,4],
                             summary(model3)$coefficients[,4]),
        bold = 0.1, caption = 'Allocation: Truncated Regression',
        caption.above = TRUE, digits = 4)
```

  + **sigma** = estimated standard deviation of the residuals

  + **model 1**
    + trunc(allocation ~ **DOSPERT risk taking** + recommended allocation)

> for a one unit increase in DOSPERT hypothetical risk taking, there is a 1.02 percentage point increase in user stock allocation 


  + **model 2**
    + trunc(allocation ~ **DOSPERT beta** + recommended allocation)

> for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 3.51 percentage point increase in user allocation

  + **model 3**
    + trunc(allocation ~ **dohmen** + recommended allocation)

> for a one unit increase in dohmen self reported risk, there is a 1.31 percentage point increase in user allocation 

\
[back to table of contents](#TOC)

#tobit/censored regression 

> technique used when true value of outcome variable is not available because of some restriction (i.e. censored)
  
  + useful for us since some users could want 110% or -10% stock allocation  

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(censReg) #http://www.ats.ucla.edu/stat/r/dae/tobit.htm

# model1 <- vglm(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, 
#                tobit(Upper = 1), data = full_data)
# 
# model2 <- vglm(allocation_stock_actual ~ beta + allocation_stock_rec, 
#                tobit(Upper = 1), data = full_data)
# 
# model3 <- vglm(allocation_stock_actual ~ dohmen + allocation_stock_rec, 
#                tobit(Upper = 1), data = full_data)

model1 <- censReg(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, 
               right = 1, data = full_data)
model2 <- censReg(allocation_stock_actual ~ beta + allocation_stock_rec, 
               right = 1, data = full_data)
model3 <- censReg(allocation_stock_actual ~ dohmen + allocation_stock_rec, 
               right = 1, data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation: Tobit (Censored) Regression',
        caption.above = TRUE, digits = 4)
```

  + **logSigma** = estimated standard deviation of the residuals in logarithmic terms 
  
  + **model 1**
    + tobit(allocation ~ **DOSPERT risk taking** + recommended allocation)

> for a one unit increase in DOSPERT hypothetical risk taking, there is a .67 percentage point increase in user stock allocation 


  + **model 2**
    + tobit(allocation ~ **DOSPERT beta** + recommended allocation)

> for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 2 percentage point increase in user allocation

  + **model 3**
    + tobit(allocation ~ **dohmen** + recommended allocation)

> for a one unit increase in dohmen self reported risk, there is a 1.49 percentage point increase in user allocation 

\
[back to table of contents](#TOC)

#arcsin transformation

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
# arc-sin transform on the proportions, which usually make them normal
  # trans.arcsine <- function(x){asin(sign(x) * sqrt(abs(x)))}
  # trans.arcsine(full_data$allocation_stock_actual)
arc_sin <- full_data %>% mutate(allocation_stock_actual = 
                                asin(allocation_stock_actual)*2/pi)

model1 <- lm(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, data = arc_sin)
model2 <- lm(allocation_stock_actual ~ beta + allocation_stock_rec, data = arc_sin)
model3 <- lm(allocation_stock_actual ~ dohmen + allocation_stock_rec, data = arc_sin)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation: Arcsin Transformation',
        caption.above = TRUE, digits = 4)
```

 + **model 1**
    + lm(allocation ~ **DOSPERT risk taking** + recommended allocation)

> for a one unit increase in DOSPERT hypothetical risk taking, there is a .73 percentage point increase in user stock allocation 


  + **model 2**
    + lm(allocation ~ **DOSPERT beta** + recommended allocation)

> for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 1.91 percentage point increase in user allocation

  + **model 3**
    + lm(allocation ~ **dohmen** + recommended allocation)

> for a one unit increase in dohmen self reported risk, there is a 1.64 percentage point increase in user allocation 

\
[back to table of contents](#TOC)

#fractional logit model 
```{r, results='asis', echo=FALSE}
# http://stackoverflow.com/questions/19893133/fractional-logit-model-in-r
# http://www.ats.ucla.edu/stat/stata/faq/proportion.htm
# Fractional Logit Models allow for inclusion of 0/1 outcomes, have correct SE's near the bounds. 

model1 <- glm(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, 
                  family=quasibinomial(link=logit), data=full_data)
model2 <- glm(allocation_stock_actual ~ beta + allocation_stock_rec, 
                  family=quasibinomial(link=logit), data=full_data)
model3 <- glm(allocation_stock_actual ~ dohmen + allocation_stock_rec, 
                  family=quasibinomial(link=logit), data=full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation: Fractional Logit',
        override.coef = list(exp(model1$coefficients), exp(model2$coefficients), exp(model3$coefficients)),
        caption.above = TRUE, digits = 4)
```

 + **model 1**
    + lm(allocation ~ **DOSPERT risk taking** + recommended allocation)

> the odds of investing in stocks increases by 3.2% for a one unit increase in DOSPERT hypothetical risk taking. 


  + **model 2**
    + lm(allocation ~ **DOSPERT beta** + recommended allocation)

> the odds of investing in stocks increases by 11.6% for a one unit increase in DOSPERT Beta value (more risk seeking behavior).

  + **model 3**
    + lm(allocation ~ **dohmen** + recommended allocation)

> the odds of investing in stocks increases by 7.1% for a one unit increase in dohmen self reported risk

\
[back to table of contents](#TOC)

#beta regression
> used for data that is bounded between 0 and 1. [see paper here](http://www.jstatsoft.org/v34/i02/paper).  

```{r, results='asis', echo=FALSE}
library(betareg) #https://stats.stackexchange.com/questions/31300/dealing-with-0-1-values-in-a-beta-regression

beta_reg <- full_data %>% 
              mutate(allocation_stock_actual = 
                     (allocation_stock_actual*645 + .5)/646) #transforming for sake of beta regression


model1 <- betareg(allocation_stock_actual ~ DOSPERT_riskTaking + allocation_stock_rec, data = beta_reg)
model2 <- betareg(allocation_stock_actual ~ beta + allocation_stock_rec, data = beta_reg)
model3 <- betareg(allocation_stock_actual ~ dohmen + allocation_stock_rec, data = beta_reg)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        override.coef = list(c(exp(model1$coefficients$mean), model1$coefficients$precision),
                             c(exp(model2$coefficients$mean), model2$coefficients$precision),
                             c(exp(model3$coefficients$mean), model3$coefficients$precision)),
        bold = 0.1, caption = 'Allocation: Beta Regression',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(allocation ~ **DOSPERT risk taking** + recommended allocation)

> the odds of investing in stocks increases by 4.75% for a one unit increase in DOSPERT hypothetical risk taking. 


  + **model 2**
    + lm(allocation ~ **DOSPERT beta** + recommended allocation)

> the odds of investing in stocks increases by 11.18% for a one unit increase in DOSPERT Beta value (more risk seeking behavior).

  + **model 3**
    + lm(allocation ~ **dohmen** + recommended allocation)

> the odds of investing in stocks increases by 11.04% for a one unit increase in dohmen self reported risk


\
[back to table of contents](#TOC)

<br>

<br>

<br>

***
***
***

