---
title: 'DOSPERT & Betterment User Data: Predicting Allocation Deviation'
author: "Salah Chafik <salah@decisionsciences.columbia.edu>"
output:
  html_document:
    theme: cosmo
    toc: yes
date: "Updated `r format(Sys.time(), '%d %B, %Y')`"
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("/git_repositories/bangla/scripts/helper_scripts/plotting_functions.R")
source("/git_repositories/betterment/scripts/helper_scripts/load_packages.R")
full_data <- readRDS("/git_repositories/betterment/data/rds_data/userANDdospert_full.RDS") %>%
              mutate(goal_type = factor(goal_type)) %>%
              rename(DOSPERT_riskTaking = finRT_sum, beta = fin_RP)
full_data$goal_type <- relevel(full_data$goal_type, ref = "RETIREMENT")
individual <- readRDS("/git_repositories/betterment/data/rds_data/userANDdospert_average.RDS")
# full_data <- read.csv("/git_repositories/betterment/data/in_process_data/pilot_8_10.csv", stringsAsFactors=F)
gamble <- read.csv("/git_repositories/betterment/data/in_process_data/pilot_8_10_gamble.csv")
invest <- read.csv("/git_repositories/betterment/data/in_process_data/pilot_8_10_invest.csv")
gamble <- left_join(full_data, gamble %>% select(finRT_sum_G:fin_RP_G, unique_id), 
                    by=c("visitor_key" = "unique_id"))
invest <- left_join(full_data, invest %>% select(finRT_sum_I:fin_RP_I, unique_id), 
                    by=c("visitor_key" = "unique_id"))
```


#Allocation Deviation 
#####using deviation (between actual and recommended allocation) as the dependent variable, we examine three significant explanatory variables: DOSPERT hypothetical risk taking, DOSPERT beta score (risk attitude), and Dohmen self reported risk

```{r, echo=FALSE, results = 'asis'}
# summary(lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)) 
# summary(lm(allocation_deviation ~ beta + allocation_stock_rec, data = full_data))
# summary(lm(allocation_deviation ~ dohmen + allocation_stock_rec, data = full_data))
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)
model2 <- lm(allocation_deviation ~ beta + allocation_stock_rec, data = full_data)
model3 <- lm(allocation_deviation ~ dohmen + allocation_stock_rec, data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: controling for recommendation',
        caption.above = TRUE, digits = 4)
```

  + **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + recommended allocation)

> **for a one unit increase in DOSPERT hypothetical risk taking, there is a .58 percentage point increase in user allocation deviation (i.e. an increase in stock allocation above the recommended amount). **

  + **model 2**
    + lm(deviation ~ **DOSPERT beta** + recommended allocation)

> **for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 2 percentage point increase in user allocation deviation**

  + **model 3**
    + lm(deviation ~ **dohmen** + recommended allocation)

> **for a one unit increase in dohmen self reported risk, there is a 1.31 percentage point increase in user allocation deviation**

##logit models 

```{r, echo=FALSE, results = 'asis'}
# summary(lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec, data = full_data)) 
# summary(lm(allocation_deviation ~ beta + allocation_stock_rec, data = full_data))
# summary(lm(allocation_deviation ~ dohmen + allocation_stock_rec, data = full_data))
model1 <- glm(factor(allocation_deviation) ~ DOSPERT_riskTaking + allocation_stock_rec, family = binomial(link = "logit"), data = full_data)
model2 <- glm(factor(allocation_deviation) ~ beta + allocation_stock_rec, family = binomial(link = "probit"), data = full_data)
model3 <- glm(allocation_deviation ~ dohmen + allocation_stock_rec,  family = binomial(link = "logit"), data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: controling for recommendation + logit model',
        caption.above = TRUE, digits = 4)
```


####scatterplot matrix (SPLOM) plot
```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=8, cache=TRUE}
ggpairs(full_data %>% select(allocation_deviation, DOSPERT_riskTaking, beta, dohmen),
        lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Allocation Deviation vs Hypothetical Risk Taking/Risk Attitude(Beta)/Dohmen") 
```

> no significant correlations between allocation deviation and three explanatory variables

\
[back to table of contents](#TOC)

#Allocation Deviation: by goal type
 
####[as seen previously](https://vlab.decisionsciences.columbia.edu/salah/user_goal_ii.html#deviations), goal type is significant as a predictive variable:
```{r, echo=FALSE}
summary(aov(allocation_deviation ~ goal_type, data=full_data))
```

```{r, echo=FALSE, results = 'asis'}
# summary(lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec + goal_type, data = full_data)) 
# summary(lm(allocation_deviation ~ beta + allocation_stock_rec + goal_type, data = full_data))
# summary(lm(allocation_deviation ~ dohmen + allocation_stock_rec + goal_type, data = full_data))
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + allocation_stock_rec + goal_type, data = full_data)
model2 <- lm(allocation_deviation ~ beta + allocation_stock_rec + goal_type, data = full_data)
model3 <- lm(allocation_deviation ~ dohmen + allocation_stock_rec + goal_type, data = full_data)

htmlreg(l = list(model1, model2, model3), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by goal type',
        caption.above = TRUE, digits = 4)
```

> **while users generally allocate more stocks than advised, they do less for `build wealth`, `emergency`, and `other` goal types, relative to the retirement goal baseline. this is true of our 3 risk taking predictors (DOSPERT risk taking, Dospert beta and dohmen) when controlling for recommended allocation. ** 

\
[back to table of contents](#TOC)

#Allocation Deviation: by DOSPERT sub-scales
```{r, echo=FALSE, results='asis'}
ss <- left_join(gamble, invest %>% select(account_id, DOSPERT_riskTaking_I = finRT_sum_I, 
                                          beta_I = fin_RP_I), by = "account_id") %>%
              rename(DOSPERT_riskTaking_G = finRT_sum_G, beta_G = fin_RP_G)
model1_ss <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + 
                                       allocation_stock_rec, data = ss)

htmlreg(l = list(model1_ss), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by DOSPERT sub-scales',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking: gamble** + **DOSPERT risk taking: invest** + recommended allocation)

> **investment is the more predictive portion of the DOSPERT scale.**

\
[back to table of contents](#TOC)

#Allocation Deviation: by Dohmen
```{r, echo=FALSE, results='asis'}
model1_d <- lm(allocation_deviation ~ DOSPERT_riskTaking + dohmen + allocation_stock_rec, data = full_data)
model2_d <- lm(allocation_deviation ~ beta +  dohmen + allocation_stock_rec, data = full_data)

htmlreg(l = list(model1_d, model2_d), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by Dohmen',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + dohmen + recommended allocation)

> no difference after inclusion of dohmen 

+ **model 2**
    + lm(deviation ~ **DOSPERT beta** + dohmen + recommended allocation)

> estimate is smaller and p-val is higher (.059 to .078) after inclusion of dohmen 

\
[back to table of contents](#TOC)

#Allocation Deviation: by DOSPERT sub-scales + Dohmen

```{r, echo=FALSE, results='asis'}
model1_dss <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + 
                                        dohmen + allocation_stock_rec, data = ss)

htmlreg(l = list(model1_dss), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: by DOSPERT sub-scales + Dohmen',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking: gamble** + **DOSPERT risk taking: invest** + dohmen + recommended allocation)

> effect size decreased and p-value increased (.001 to .008) by inclusion of dohmen

\
[back to table of contents](#TOC)

#Allocation Deviation: combining risk taking predictors

```{r, echo=FALSE, results='asis'}
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec, data = full_data) 
model2 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type, data = full_data) 
model3 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + DOSPERT_riskTaking*goal_type, 
                                    data = full_data) #Drt
model4 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + beta*goal_type, 
                                    data = full_data) #beta
model5 <- lm(allocation_deviation ~ DOSPERT_riskTaking + beta + dohmen +  
                                    allocation_stock_rec + goal_type + dohmen*goal_type, 
                                    data = full_data) #dohmen
htmlreg(l = list(model1, model2, model3, model4, model5), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: Combining Risk Taking Predictors',
        caption.above = TRUE, digits = 4)
```

+ **model 1**
    + lm(deviation ~ **DOSPERT risk taking** + **DOSPERT beta** + **dohmen** + recommended allocation)

> **for a one unit increase in DOSPERT hypothetical risk taking, there is a .39 percentage point increase in user allocation deviation (i.e. an increase in stock allocation above the recommended amount).**

+ **model 2**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + **goal type** + recommended allocation)

> **with the inclusion of goal type in the model, the same significant relationship holds true for DOSPERT hypothetical risk taking as in Model 1. in addition, for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 1.76 percentage point increase in user allocation deviation. finally, users allocate less stocks for `build wealth`, `emergency` and `other` goal types, relative to the retirement goal baseline.**

+ **model 3**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **DOSPERT risk taking**X**goal_type**  + recommended allocation)
  
> with the inclusion of Dospert risk taking and goal type interaction, the same significant relationship holds true for DOSPERT beta as in Model 2.

+ **model 4**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **DOSPERT beta**X**goal_type**  + recommended allocation)

> **with the inclusion of Dospert beta and goal type interaction, the same significant relationship holds true for DOSPERT risk taking as in Models 1 + 2. in addition, for a one unit increase in DOSPERT Beta value (more risk seeking behavior), there is a 10.72/12.54 percentage point increase in user allocation deviation for `house`/`purchase` goal types.** finally, users allocate less stocks for `other` goal type, relative to the retirement goal baseline.

+ **model 5**
    + lm(deviation ~ DOSPERT risk taking + DOSPERT beta + dohmen + goal type + **dohmen**X**goal_type**  + recommended allocation)

> with the inclusion of dohmen and goal type interaction, the same significant relationship holds true for DOSPERT risk taking as in Models 1 + 2 + 4. the same significant holds true for DOSPERT beta from Models 2 + 3. in addition, users allocate less stocks for `education` goal type, relative to the retirement goal baseline.

\
[back to table of contents](#TOC)


```{r, echo=FALSE, results='asis', eval=FALSE}
#Allocation Deviation: combining risk taking predictors (sub-scales)
model1 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec, data = ss) 
model2 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + 
                                    DOSPERT_riskTaking_G*goal_type + DOSPERT_riskTaking_I*goal_type, 
                                    data = ss) #Drt
model3 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + beta*goal_type, 
                                    data = ss) #beta
model4 <- lm(allocation_deviation ~ DOSPERT_riskTaking_G + DOSPERT_riskTaking_I + beta + dohmen +  
                                    allocation_stock_rec+ goal_type + dohmen*goal_type, 
                                    data = ss) #dohmen
htmlreg(l = list(model1, model2, model3, model4), center = TRUE, stars = c(0.001, 0.01, 0.05, 0.1),
        bold = 0.1, caption = 'Allocation Deviation: Combining Risk Taking Predictors (DOSPERT sub-scales)',
        caption.above = TRUE, digits = 4)
\
[back to table of contents](#TOC)
```



```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=14, fig.height=7, eval=FALSE}
#Actual Allocation? 
ggpairs(full_data %>% select(allocation_stock_actual, DOSPERT_riskTaking),
        lower=list(continuous="smooth", params=c(colour="blue")),
        diag=list(continuous="density", params=c(colour="blue")), 
        upper=list(params=list(corSize=5)), axisLabels='show',
        title = "Actual Stock Allocation vs Hypothetical Risk Taking SPLOM Plot") 

> there is a weak (although strongly significant) positive correlation of .11 between actual stock allocation and DOSPERT hypothetical risk taking 

```




